\ifdefined\MAINTEXT \else Supplemental Material: \\ \fi Low-cost linear-scaling \emph{ab initio} molecular dynamics for weakly-interacting systems

### 2018-03-06 almo-langevin.tex
Small edits
Went through and edited, made small edits in grammar.

In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors. 
%~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}. 
Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016} explore the natural locality of the one-electron density matrix (DM). 
However, the DM DFT becomes advantageous only for impractically large systems when accurate multifunction basis sets are used~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:almo-ls}.
%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
This issue is rectified in optimal-basis DM methods~\cite{Skylaris2005, Nakata2015, Mohr2015} that contract large basis sets into a small number of new localized functions and then optimize the DM in the contracted basis. 
%Later: Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} 
%While most large-scale DFT calculations are performed using this class of methods
Despite becoming the most popular approach to LS DFT the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
%
From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved with significantly fewer variables. 
Advantages of the orbitals-only LS DFT are especially pronounced in accurate calculations that require many basis functions per atom.
%From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%
% direct optimization of such compact localized nonorthogonal orbitals
%
%To obviate the DM optimization, another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:weitao-yang-2013,a:ls-tsuchida-aomm} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-rev-1999,a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:weitao-yang-2013,a:ls-tsuchida-aomm}.

@@

Thus, despite impressive progress of LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign} the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}. 
%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign}
On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see ALMO SCF method in Ref.~\citenum{a:almo-ls}.

ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and, if the exchange-correlation (XC) functional includes them, dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance in finite-gap materials~\cite{a:ls-rev-1999}, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems. %with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{Kuhne2013}. 
Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces:
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise: %with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
% 
The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at a specified temperature $T$~\cite{a:Kubo-1986}. 
In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled. 
%Increasing $\gamma$ leads to faster equilibration but more distorted results for dynamical properties.

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t')
\end{align}
%
This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces:
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t)
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD.

@@

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
\caption{\label{fig:dynproperties} 
Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) RDF, 
(b) kinetic energy distribution (the black curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) IR spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

@@

The \emph{reference} forces were calculated with fully delocalized electrons using tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
In ALMO AIMD, element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR). This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the reference radial distribution function (RDF) perfectly in Monte-Carlo simulations~\cite{a:almo-ls}. 
%
To check the ability of the $R^{\Delta}(t)$ term to compensate for imperfections in ALMO forces we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.

@@

To test the accuracy of ALMO AIMD we used trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, RDF, and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR) and from the reference trajectory. 
The diffusion coefficients $D_{\text{OT}}=\SI{1.9\pm0.9E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{1.8\pm0.4E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement. The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals. These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a) the compensating $R^{\Delta}(t)$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly. We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$. Neither we were able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.

@@

% RZK longer article: averaging is done over 10 time steps 
To demonstrate computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
%In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}. 
It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions -- a particularly challenging case for DM-based LS methods.
ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$ even for medium-size systems. While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules -- length scales routinely accessible with AIMD today.

@@

Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution: LS is retained for a wide range of systems and compute cores. 
We were able to successfully simulate systems as large as $\sim10^{5}$ atoms 
%(32768 water molecules) 
within reasonable wall-clock time using only moderate number of compute cores -- an impressive feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources. It indicates that ALMO AIMD can extend the range of routine simulations to $\sim10^4$ atoms on modern HPC platforms.

@@

% systematically improvable approximation to DFT-based AIMD.
To summarize, we demonstrated, for the first time, that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM. %for weakly-interacting systems. 
High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces. 
By obviating the optimization of the DM the method remains remarkably efficient even with large localized basis sets. 
%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible time and length scales. 
The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces, nuclei) making completely new phenomena accessible to AIMD. 
Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.

>>>>

In all LS DFT methods, the \emph{delocalized} eigenstates of the effective KS Hamiltonian must be replaced with an alternative set of \emph{local} electronic descriptors. 
%~\footnote{The delocalization of KS orbitals over all atoms is the main reason for the steep growth of the cost of KS DFT: since both the number of orbitals and the number of localization regions (e.g. atomic basis functions) increase linearly with the number of atoms, the total number of the electronic degrees of freedom grows quadratically.}. 
Most LS methods~\cite{a:ls-rev-1999, a:ls-rev-2012, Kussmann2013, Aarons2016} explore the natural locality of the one-electron density matrix (DM). 
However, the DM DFT becomes advantageous only for impractically large systems when accurate multifunction basis sets are used~\cite{a:ls-dm-sign, Arita2014, a:ls-rev-2012, a:almo-ls}.
%Therefore, the application of DM-based LS methods have been restricted to minimal-basis tight-binding problems~\cite{Richters2014, more}. 
This issue is rectified in optimal-basis DM methods~\cite{Skylaris2005, Nakata2015, Mohr2015} that contract large basis sets into a small number of new localized functions and then optimize the DM in the contracted basis. 
%Later: Although such methods have been successfully used for the evaluation of accurate DFT energies of large static systems~\cite{a:ls-onetep-2009,a:ls-conquest-2010,a:ls-onetep-2010-app1,a:ls-rev-2012} 
%While most large-scale DFT calculations are performed using this class of methods
Despite becoming the most popular approach to LS DFT, the efficiency of these methods is hampered by the costly optimization of both the contracted orbitals and the DM~\cite{a:ls-onetep-2003}. 
%
From this point of view, a direct variation of molecular orbitals that are strictly localized within predefined regions is preferable because LS can be achieved with significantly fewer variables. 
Advantages of the orbitals-only LS DFT are especially pronounced in accurate calculations that require many basis functions per atom.
%From this point of view, it is desirable to obviate the DM optimization and use another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%
% direct optimization of such compact localized nonorthogonal orbitals
%
%To obviate the DM optimization, another set of local electronic descriptors can be constructed by enforcing strict locality on one-electron orbitals and, necessarily, by relaxing the orthogonality constraints imposed on them. 
%From the computational point of view, LS methods based on the direct optimization of such compact localized nonorthogonal orbitals~\cite{a:ls-galli-parrinello-1992,a:ls-mauri-galli-car-1993,a:ls-ordejon-1993,a:ls-mauri-galli-1994,a:ls-ordejon-1995,a:ls-kim-mauri-galli-1995,a:ls-fattebert-2004,a:ls-fattebert-2006,a:burger-yang-2008} are promising since they deal with fewer variational degrees of freedom than the DM methods. 
Unfortunately, the development of promising orbital-based LS methods has been all but abandoned~\cite{a:weitao-yang-2013,a:ls-tsuchida-aomm} because of the inherently difficult optimization of localized orbitals~\cite{a:ls-rev-1999,a:ls-mauri-galli-car-1993,a:ls-ordejon-1995,a:ls-fattebert-2004,a:weitao-yang-2013,a:ls-tsuchida-aomm}.

@@

Thus, despite impressive progress of the LS description of the electronic and atomic structure of large static systems~\cite{Bowler2010,a:ls-dm-sign}, the high computational overhead of existing LS methods restrict their use in dynamical simulations to very short \emph{time scales}, systems of low dimensions, and low-quality minimal basis sets~\cite{a:ls-dm-sign, Otsuka2016, Hine2011, Bowler2010}. 
%RZK: Although such methods have been successfully used for the evaluation of accurate DFT energies [electronic structure and atomic configurations] of large static systems~\cite{Bowler2010,a:ls-dm-sign}
On typical length and time scales required in practical and accurate AIMD simulations, LS DFT still cannot compete with the straightforward low-cost cubically-scaling KS DFT.

@@

The new AIMD method utilizes a recently developed LS DFT~\cite{a:almo-ls} based on absolutely localized molecular orbitals (\mbox{ALMOs}). 
Unlike delocalized KS orbitals, each \mbox{ALMO} has its own \emph{localization center} and a predefined \emph{localization radius} $R_{c}$ that typically includes nearby atoms or molecules~\cite{a:stoll,a:almo-ls}. 
In the current implementation, a localization center is defined as a set of all Gaussian atomic orbitals of one molecule. 
However, the approach can use other local and nonlocal basis sets~\cite{a:galli_loc, Lin2012}. 
The key feature of ALMO DFT is that its one-electron wavefunctions are constructed in a two-stage self-consistent-field (SCF) procedure~\cite{a:almo-ls} to circumvent the problem of the sluggish variational optimization emphasized above. 
In the first stage, ALMOs are constrained to their localization centers~\cite{a:khal} whereas, in the second stage, ALMOs are relaxed to allow delocalization onto the neighbor molecules within their localization radius $R_{c}$. 
To achieve a robust optimization in the problematic second stage, it is important to keep the delocalization component of the trial wavefunction orthogonal to the fixed orbitals obtained in the first stage. 
For mathematical details, see the ALMO SCF method in Ref.~\citenum{a:almo-ls}.

ALMO constraints imposed by $R_c$ prohibit electron density transfer between distant molecules, but retain all other types of interaction such as long-range electrostatic, exchange, polarization, and---if the exchange-correlation (XC) functional includes them---dispersion interactions~\cite{a:theeda}. 
Since the importance of electron transfer decays exponentially with distance in finite-gap materials~\cite{a:ls-rev-1999}, the \mbox{ALMO} approximation is expected to provide a natural and accurate representation of the electronic structure of molecular systems. %with just a few local variables. 
Because of the greatly reduced number of electronic descriptors and the robust optimization, the computational complexity of ALMO DFT grows linearly with the number of molecules while its computational overhead remains very low. These features make ALMO DFT a promising method for accurate AIMD simulations of large molecular systems.

@@

Our approach to compensate for the missing $\delta f_{i\alpha}(t)$ term is inspired by the methodology introduced into AIMD by Krajewski \emph{et al.}~\cite{Krajewski}, formalized by K\"uhne \emph{et al.}~\cite{a:2ndcpmd} and rationalized by Dai \emph{et al.}~\cite{a:langevin-why} before becoming informally known as the second generation Car-Parrinello molecular dynamics~\cite{Kuhne2013}. 
Adopting the principle of Refs.~\onlinecite{Krajewski} and \onlinecite{a:2ndcpmd}, ALMO AIMD is chosen to be governed by the Langevin equation of motion that can be written in terms of the unknown reference forces
%
\begin{align}
\label{eq:langevin}
m_i \ddot{r}_{i\alpha} = f^{\text{KS}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma}_{i\alpha} (t),
\end{align}
%
where $m_i$ is the mass of atom $i$, $r_{i\alpha}$ is its position along dimension $\alpha$, $\gamma$ is the Langevin scaling factor, and $R^{\gamma}_{i\alpha} (t)$ is the stochastic force represented by a zero-mean white Gaussian noise %with the zero mean and the strength related to the damping term through $\gamma$:
%
\begin{align}
\label{eq:stochastic}
\langle R^{\gamma}_{i\alpha} (t) \rangle &= 0, \\
\langle R^{\gamma}_{i\alpha} (t)  R^{\gamma}_{j\beta} (t') \rangle &= 2 k_B T \gamma m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t').
\end{align}
% 
The last relation means that, for any value of $\gamma$, the damping and stochastic terms are in perfect balance and trajectories generated with Eq.~(\ref{eq:langevin}) will sample the canonical ensemble at a specified temperature $T$~\cite{a:Kubo-1986}. 
In the limit $\gamma \rightarrow 0$, the Newton equation is recovered and the microcanonical ensemble is sampled. 
%Increasing $\gamma$ leads to faster equilibration but more distorted results for dynamical properties.

The main assumption of ALMO AIMD is that the error in the ALMO forces is well approximated by Gaussian noise $R^{\Delta}_{i\alpha} (t)$:
%
\begin{align}
\label{eq:assumption}
\delta f_{i\alpha}(t) = R^{\Delta}_{i\alpha} (t)
\end{align}
%
that obeys
%
\begin{align}
\label{eq:stochastic2}
\langle R^{\Delta}_{i\alpha} (t) \rangle &= 0, \\
\label{eq:stochastic3}
\langle R^{\Delta}_{i\alpha} (t)  R^{\Delta}_{j\beta} (t') \rangle &= 2 k_B T \Delta m_i \delta_{ij} \delta_{\alpha\beta} \delta(t-t') .
\end{align}
%
This assumption, shown to be well justified, allows us to rewrite the Langevin equation using the ALMO forces
%
\begin{align}
\label{eq:langevin2}
m_i \ddot{r}_{i\alpha} = f^{\text{ALMO}}_{i\alpha}(t) - \gamma m_i \dot{r}_{i\alpha} + R^{\gamma + \Delta}_{i\alpha} (t),
\end{align}
%
where the two stochastic terms are combined into one $R^{\gamma + \Delta}_{i\alpha} = R^{\gamma}_{i\alpha} + R^{\Delta}_{i\alpha}$. 
The only missing piece in the modified Langevin equation is the value of $\Delta$, which describes the strength of the newly introduced stochastic term. 
This term compensates for imperfections in ALMO forces and must be adjusted to re-balance the damping and stochastic components in ALMO AIMD.

@@

\begin{figure}
\includegraphics[trim={1.3cm 0.1cm 3.3cm 1.3cm},clip,width=8.6cm]{Dynamical_Data_Tiled.eps}
\caption{\label{fig:dynproperties} 
Calculated properties of water using ALMO AIMD with $\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR (red line) and fully converged OT reference (black line).
(a) RDF, 
(b) kinetic energy distribution (the gray curve shows the theoretical Maxwell-Boltzmann distribution), 
(c) IR spectrum.
%Calculations were performed with the PBE $\mathrm{E_{XC}}$ functional and TZV2P basis set, using 125 water molecules for 10 ps.
}
\end{figure}

@@

The \emph{reference} forces were calculated with fully delocalized electrons using the tightly converged, $\epsilon_{\text{SCF}}=10^{-6}$~a.u., orbital transformation (OT) method~\cite{a:ot}. 
In ALMO AIMD, the element-specific cutoff radius for electron delocalization $R_c$ was set to 1.6 in units of the elements' van der Waals radii (vdWR). This localization radius includes approximately two coordination shells of an average water molecule and was shown to reproduce the reference radial distribution function (RDF) perfectly in Monte-Carlo simulations~\cite{a:almo-ls}. 
%
To check the ability of the $R^{\Delta}(t)$ term to compensate for imperfections in ALMO forces, we varied $\epsilon_{\text{SCF}}$ between tight $10^{-6}$~a.u. and loose $10^{-2}$~a.u.

@@

To test the accuracy of ALMO AIMD we used the trajectory analyzer TRAVIS~\cite{a:travis-main} to compute the infrared (IR) spectrum, RDF, and diffusion coefficient of liquid water from both the ALMO trajectory ($\epsilon_{\text{SCF}} = 10^{-2}$~a.u. and $R_{c} = 1.6$ vdWR) and from the reference trajectory. 
The diffusion coefficients $D_{\text{OT}}=\SI{1.9\pm0.9E-10}{\m^{2}\cdot\s^{-1}}$ and $D_{\text{ALMO}}=\SI{1.8\pm0.4E-10}{\m^{2}\cdot\s^{-1}}$ and RDFs (Figure~\ref{fig:dynproperties}a) are in good agreement. The quality of the ALMO IR spectrum (Figure~\ref{fig:dynproperties}c) is good despite minor errors in the intensity of the OH stretching mode, which is sensitive to the precise positions of the centers of localized orbitals. These stringent tests show that despite noticeable errors in the ALMO forces (Figure \ref{fig:randomforce}a), the compensating $R^{\Delta}(t)$ term in the modified Langevin equation makes it possible to recover atomic dynamics properly. We would like to note that ALMO AIMD could not be stabilized with $\Delta=0$. Neither were we able to find any values of $\Delta$ that stabilize trajectories generated using perturbative versions of ALMO DFT~\cite{a:almo-ls}.

@@

% RZK longer article: averaging is done over 10 time steps 
To demonstrate the computational efficiency of ALMO AIMD, we compared the average wall-time per MD step for a variety of methods in Figure \ref{fig:strongscaling_log}.
%In addition to the fully converged OT reference, we included timing for the second generation Car-Parrinello method designed to decrease the computational prefactor of the cubically-scaling DFT~\cite{a:2ndcpmd}. 
It is important to emphasize that the comparison is performed for a three-dimensional condensed phase system described with an accurate triple-$\zeta$ basis set with polarized functions---a particularly challenging case for DM-based LS methods.
ALMO AIMD shows clear LS behavior for all values of $\epsilon_{\text{SCF}}$, even for medium-size systems. While the second generation Car-Parrinello method decreases the computational overhead of the cubically-scaling AIMD for small systems~\cite{a:2ndcpmd}, ALMO AIMD exploits the modified Langevin concept to substantially reduce the simulation cost for systems of all sizes.
The crossover point between ALMO AIMD and cubically scaling methods lies in the region of 256 molecules---length scales routinely accessible with AIMD today.

@@

Weak scaling benchmarks for very large systems show (Figure~\ref{fig:weakscaling}) that localized orbitals are naturally suited for parallel execution: LS is retained for a wide range of systems and compute cores. 
We were able to successfully simulate systems as large as $\sim10^{5}$ atoms 
%(32768 water molecules) 
within reasonable wall-clock time using only moderate number of compute cores---an impressive feat for AIMD considering that accurate molecular orbitals and the idempotent DM are computed on each step.
The horizontal line in Figure~\ref{fig:weakscaling} is shown as a rough guide to time and length scales accessible in a fixed wall-clock time given various computational resources. It indicates that ALMO AIMD can extend the range of routine simulations to $\sim10^4$ atoms on modern HPC platforms.

@@

% systematically improvable approximation to DFT-based AIMD.
To summarize, we demonstrated---for the first time---that compact localized orbitals can be utilized to perform accurate and efficient LS AIMD without concomitant optimization of the DM. %for weakly-interacting systems. 
High efficiency of the presented method is achieved without sacrificing accuracy with a combination of two techniques: (1) on-the-fly calculation of approximate forces without lengthy self-consistent optimization of localized orbitals and (2) integration of a modified Langevin equation of motion that is fine-tuned to retain stable dynamics even with imperfect forces. 
By obviating the optimization of the DM, the method remains remarkably efficient even with large localized basis sets. 
%A remarkable feature of the method is that it remains efficient for challenging condensed phase systems even with large localized basis sets. 
Using liquid water as an example, we showed that the new approach enables simulations of molecular systems on previously inaccessible time and length scales. 
The developed method will have a significant impact on modeling of complex molecular systems (e.g. interfaces or nuclei) making completely new phenomena accessible to AIMD. 
Generalization of the methodology to systems of strongly interacting atoms (e.g. covalent crystals) is underway.
